---
title: "FinalProject_analysis"
output: pdf_document
date: "2022-12-12"
---

```{r}
if(!require(tidyverse)) install.packages("tidyverse")
if(!require(sqldf)) install.packages("sqldf")

library(tidyverse)
library(sqldf)
```

Read the CSV files
```{r}
players <- read.csv("/Users/devyanimardia/Downloads/datadavis-nba-salaries/players.csv") 
salaries <- read.csv("/Users/devyanimardia/Downloads/datadavis-nba-salaries/salaries_1985to2018.csv") 
```

Joining both datasets with player_id into salaries dataset
```{r}
salaries <- salaries %>% inner_join(players, by = c("player_id" = "X_id"))
```

Age, Height and Weight changes to make it easily usable
```{r}
#Age at end of season

salaries$birthDate <- sub(",","", salaries$birthDate)

salaries$birthDate <- as.Date(salaries$birthDate,
  format = "%B %d %Y")

salaries$Current_age = as.numeric(difftime(Sys.Date(),salaries$birthDate, units = "weeks"))/52.25

salaries$AgeatSeasonEnd= as.integer(salaries$Current_age - (2022- salaries$season_end))
#Weight

salaries$weight <- as.numeric(sub("lb","", salaries$weight))

#Height
salaries$height = as.character(sub("-",".", salaries$height))

salaries <- salaries %>% separate(height, c('Feet', 'Inch'))

salaries$heightincm <- as.numeric(salaries$Feet)*30.48 + as.numeric(salaries$Inch)*2.54


head(salaries)
```
Converting to numeric for better usage
```{r}
salaries$career_FG3. <- as.numeric(salaries$career_FG3.)
salaries$career_FG. <- as.numeric(salaries$career_FG.)
salaries$career_FT. <- as.numeric(salaries$career_FT.)
salaries$career_PER <- as.numeric(salaries$career_PER)
salaries$career_TRB <- as.numeric(salaries$career_TRB)
salaries$career_WS <- as.numeric(salaries$career_WS)
salaries$career_eFG. <- as.numeric(salaries$career_eFG.)
salaries$heightincm <- as.numeric(salaries$heightincm)


```

Variable name changes and slicing the salaries csv to only get the variables important for us

```{r}
experimentwithseason <- salaries[c("player_id","heightincm", "weight","season_end", "AgeatSeasonEnd","team","career_AST","career_PTS","career_FG.","career_FG3.","position","shoots","salary","career_FT.","career_G","career_PER","career_TRB","career_WS","career_eFG.")] 


names(experimentwithseason) = c("id","height", "weight","season_end", "AgeatSeasonEnd","team","Assits","Points","FieldGoals","FieldGoals3","position","shoots","salary","FreeThrows","Games","PlayerEfficiencyRating","TotalRebounds","WinShares","effectiveFieldGoal")

```

Factorize categorical variables
```{r}
experimentwithseason$team.f <- factor(experimentwithseason$team)
is.factor(experimentwithseason$team.f)

experimentwithseason$shoots.f <- factor(experimentwithseason$shoots)
is.factor(experimentwithseason$shoots.f)

experimentwithseason$position.f <- factor(experimentwithseason$position)
is.factor(experimentwithseason$position.f)

```
Creating 2015 dataframe and removing unwanted columns
```{r}

df_2015 = subset(experimentwithseason, season_end == 2015)


df_2015 = subset(df_2015, season_end == 2015,select = -c(team,shoots,position))

names(df_2015) = c("id","height", "weight","season_end", "AgeatSeasonEnd","Assits","Points","FieldGoals","FieldGoals3","salary","FreeThrows","Games","PlayerEfficiencyRating","TotalRebounds","WinShares","effectiveFieldGoal","team","shoots","position")

df_2015 <- sqldf('select weight, height, AgeatSeasonEnd, team, Assits, FieldGoals, FieldGoals3, position, shoots, FreeThrows, Games, PlayerEfficiencyRating, TotalRebounds, WinShares, effectiveFieldGoal, Points, salary from experimentwithseason where season_end == 2015  group by id')






str(df_2015)

head(df_2015)
```

Fitting an initial model with no changes 

```{r}

str(df_2015)
model1 = lm(salary ~ ., data=df_2015)
summary(model1)
plot(model1)

```
```{r}
library(MASS)
lambda = 0.303
box_cox = boxcox(model1)
#lambda= box_cox$x[which.max(box_cox$y)]
#lambda
modeltest1 = lm(((salary^lambda-1)/lambda)~., df_2015)
summary(modeltest1)
plot(modeltest1)

```
Forward Selection
```{r}
n = nrow(df_2015)
out.null=lm(salary ~1, df_2015)
out.full=lm(salary~.,df_2015)
full=formula(lm( salary~., df_2015))
#out.forward= step(out.null,scope=list(lower=~1,upper=full), k= 2,direction="forward",trace= TRUE)
```

```{r}
#out.forward= step(out.null,scope=list(lower=~1,upper=full), k= log(n),direction="forward",trace= TRUE)


```
Backward Elimination
```{r}
#out.backward=step(out.full ,scope=list(lower=~1,upper=full),direction="backward",trace=TRUE ,k=log(n))

```

```{r}
#out.backward=step(out.full ,scope=list(lower=~1,upper=full),direction="backward",trace=TRUE ,k= 2)

  
```
```{r}
model_3_forwardsel_param = lm(salary ~ WinShares + Points + weight + Assits, data=df_2015)
summary(model_3_forwardsel_param)
plot(model_3_forwardsel_param)

model_6_back_param = lm(salary ~ height + weight + AgeatSeasonEnd + Assits + Points + 
    FieldGoals + FieldGoals3 + FreeThrows + Games + PlayerEfficiencyRating + 
    TotalRebounds + WinShares + effectiveFieldGoal, data=df_2015)
summary(model_6_back_param)
plot(model_6_back_param)

```

```{r}
#The frequent positions with the positions which are used in categorical variables
df_2015_frequentpos <- df_2015 %>% filter(position %in% c('Point Guard','Center','Shooting Guard','Power Forward','Small Forward'))

```

```{r}
#Best model

lambda=0.222
model_9 = lm(((salary^lambda-1)/lambda)~ ., data = df_2015_frequentpos)
plot(model_9)
summary(model_9)
```
```{r}
z = experimentwithseason %>% separate_rows(position, sep = " and ") %>% filter(season_end == 2015)

df_2015_2 = subset(z, season_end == 2015,select = -c(id,team,shoots,position))

names(df_2015_2) = c("height", "weight","season_end", "age_atendofseason","Assits","Points","FieldGoals","FieldGoals3","salary","FreeThrows","Games","PlayerEfficiencyRating","TotalRebounds","WinShares","effectiveFieldGoal","team","shoots","position")

```

```{r}
model12 = lm(salary ~ .,df_2015_2)
summary(model12)
plot(model12)

```
```{r}
box_cox = boxcox(model1)
lambda= box_cox$x[which.max(box_cox$y)]
lambda
modeltest2 = lm(((salary^lambda-1)/lambda)~., df_2015_2)
summary(modeltest2)
plot(modeltest2)
```
```{r}
m = sqldf('select height, avg(salary) as meansal, season_end from experimentwithseason group by height, season_end order by meansal, season_end')

m = sqldf('select height, avg(salary) as meansal, season_end from experimentwithseason where height> 200 group by height, season_end order by meansal, season_end')

ggplot(m, aes(height, meansal)) + geom_col()
ggplot(m, aes(season_end, meansal, fill = height)) + geom_col()

```
```{r}
expafter006 = subset(experimentwithseason, season_end > 2000,select = -c(id))

expafter006 %>% group_by(team) %>% count() %>% arrange(desc(n)) %>% ggplot(aes(n, team)) + geom_col() 

expafter006 %>% group_by(season_end) %>% count() %>% ggplot(aes(season_end, n)) + geom_col()

```
```{r}
expafter006 %>% group_by(season_end, team) %>% count() %>% ggplot(aes(season_end, n, fill = team))+ geom_col(col = "black", size = 0.2) 

```
```{r}
shapiro.test(residuals(model_9))
shapiro.test(residuals(model1))
```

